Flattering Advice: Avoiding Disappointment as a Driver of Gender Discrimination

### 2024-01-14 manuscript/manuscript.Rmd
Edits up to the beginning of Methods of Study 2.

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study1_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study1_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study1_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.") %>%
  kableExtra::kable_styling(latex_option = "scale_down")

@@

## Discussion
When the quality of a decision is not dependent on someone's expectations,
advice should not be moved by whether an adviser is aware of the advisees
expectations. However, consistent with our account that advisers consider the
belief utility of the advisee and want to flatter, rather than disappoint them,
we find that displaying expectations to advisers does matter.
Specifically, we found that men underestimate their score on a mathematics test
less than do women. When these expectations are displayed to advisers, they
are more likely to tell men to compete against the High Performer group.
Notably, this is bad advice: men's expected earnings decreases if their
expectations are shown to advisers.

@@

We then conduct a three-stage experiment to determine why advisers give
flattering advice. First, we ask a group of advisees to upload selfies
and predict their rank in a group of ten people of the same gender based
on physical attractiveness. Next, we recruit and ask advisers to rank
selfies of the opposite gender and recommend which rank the participant
ranked 7th should bet on, with the bonus being contingent on whether the
participant guessed accurately or found the adviser likable. Finally, we
show the advisees the advice they have received and ask them to make
their final prediction about their rank and evaluate the advice giver in
terms of likability, warmth, and trustworthiness. Our main hypothesis is
that advisers who want to be liked by advisees would recommend a higher
rank.

@@

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional.") %>%
  kableExtra::kable_styling(latex_option = "scale_down")

>>>>

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study1_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study1_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study1_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.")

@@

## Discussion
When the quality of a decision is not dependent on someone's expectations,
advice should not be moved by whether an adviser is aware of the advisees
expectations. However, consistent with our account that advisers consider the
belief utility of the advisee and want to flatter, rather than disappoint them,
we find that displaying expectations to advisers does matter.
Specifically, we found that men underestimate their score on a mathematics test
less than do women. When these expectations are displayed to advisers, they
are more likely to tell men to compete against the High Performer group.
Notably, this is bad advice: men whose advisers observed their expectations
got worse advice as a result.

@@

<!-- David: Just commenting on what I tried to do here... a bit of a transition -->
<!-- for why this study follows from the last. -->
So far, we have shown that displaying expectations can induce gender
differences in advice given. Specifically, advisers were more likely to
recommend competing against a group of High Performers to men who had higher
expectations about their performance. We now test directly whether this 
is driven by interpersonal concerns. Specifically, we examine whether 
more flattering advice is the result of a desire to be liked and whether
advisees indeed do like advisers more when they give flattering advice.

We conducted a three-stage experiment in which we experimentally manipulated
the incentives for advicers. We began by inviting a group of advisees to upload 
photos of themselves ("selfies") and informed them that they would be rated
on their attractiveness. We grouped them with nine other participants of the
same gender and recruited participants of the opposing gender to rank them from
most to least attractive and to provide advice. Specifically, we asked them
to advise the participant they ranked as the 7th most attractive (4th least attractive)
on what rank they should bet they were ranked by a larger group of raters.
Advisers were randomly assigned to two treatments, receiving a bonus payment
either if the advisee guessed their rank accurately or if the advisee evaluated
the adviser as likeable as measured by an unincentivized scale response.
We hypothesize that advisers who want to be liked by advisees will recommend 
that they bet on a lower rank, i.e. that they are more attractive.

@@

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional.")
